---
title: "CEV briefing QA - QCOVID analysis"
author: "Sebastien Peytrignet"
date: "6/7/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and set-up

```{r setup.B, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
##############################################
################### SETUP ####################
##############################################

#Load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,stringr,ggplot2,DescTools,data.table,kableExtra,
               tibble,pbapply,here,tidyverse,readxl,easycsv)

#Clean up the global environment
rm(list = ls())

#Set directory where inputs are saved
rawdatadir <- "M:/Analytics/Networked Data Lab/Shielding/"
briefonedrivedir <- "C:/Users/SebastienP/OneDrive - The Health Foundation/Briefing chart data/"
gitdir <- "M:/Analytics/Sebastien/GitHub/NHSD-QCOVID-analysis/"

```

## Create new look-up: IMD scores at the CCG level using IMD scores at the LSOA level and using population to aggregate upwards

```{r new.CCG.measures, eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
#######################################################################
################### LOOKUP: LOCAL AUTHORITY TO HIGHER #################
#######################################################################

LSOA_to_higher_geo <- fread(paste0(rawdatadir,"Other data/Lookups/LSOA_(2011)_to_Clinical_Commissioning_Groups_to_Sustainability_and_Transformation_Partnerships_(April_2020)_Lookup_in_England.csv"), header=TRUE, sep=",", check.names=T) %>%
  select(.,LSOA11CD,CCG20CD,CCG20CDH,CCG20NM,STP20CD,STP20NM)

length(unique(LSOA_to_higher_geo$CCG20CD)) #135 CCGs in 2020

##############################################
################### POPULATION ###############
##############################################

pop_by_LSOA <- read_excel(paste0(rawdatadir,"Other data/Mid-year population estimates/LSOA/SAPE22DT2-mid-2019-lsoa-syoa-estimates-unformatted.xlsx"),
                        sheet = "Mid-2019 Persons",skip=4) %>%
  select(.,`LSOA Code`,`LSOA Name`,`All Ages`) %>%
  dplyr::rename(.,LSOA11CD=`LSOA Code`,LSOA11NM=`LSOA Name`,pop19=`All Ages`) %>%
  filter(.,!is.na(LSOA11CD))

#Join with geo-indicators to match English local authorities to regions

pop_by_LSOA <- left_join(pop_by_LSOA,LSOA_to_higher_geo,
                              by=c("LSOA11CD"="LSOA11CD")) %>%
  filter(.,!is.na(CCG20CD))

rm(LSOA_to_higher_geo)

#Compute population by CCG

pop_by_LSOA_agg <- pop_by_LSOA %>%
  group_by(CCG20CD,CCG20CDH) %>%
  summarise(pop19=sum(pop19,na.rm=TRUE)) %>%
  ungroup()

##############################################
################### DEPRIVATION ##############
##############################################

IMD2019_LSOA <- fread(paste0(rawdatadir,"Other data/IMD/LSOA/Index_of_Multiple_Deprivation_(December_2019)_Lookup_in_England.csv"), header=TRUE, sep=",", check.names=T) %>%
  select(.,LSOA11CD,IMD19)

#Join with population and lookup data

IMD2019_LSOA <- left_join(IMD2019_LSOA,pop_by_LSOA,
                         by=c("LSOA11CD"="LSOA11CD"))

#Weighted average score by CCG and new quantiles

IMD2019_LSOA_agg <- IMD2019_LSOA %>%
  group_by(CCG20CD,CCG20CDH) %>%
  summarise(IMD19_CCG_wgt=weighted.mean(IMD19, pop19)) %>%
  ungroup() %>%
  mutate(.,IMD19_decile=ntile(IMD19_CCG_wgt, 10),
         IMD19_quintile=ntile(IMD19_CCG_wgt, 5))

##################################################
################### SAVE NEW LOOKUP ##############
##################################################

fwrite(pop_by_LSOA_agg, file = paste0(rawdatadir,"Other data/Mid-year population estimates/CCG/pop_by_LSOA_agg.csv"), sep = ",")
fwrite(IMD2019_LSOA_agg, file = paste0(rawdatadir,"Other data/Lookups/CCG_to_IMD_2020_lookup.csv"), sep = ",")
```

## Import population by CCG, IMD scores by CCG and Shielded Patient list by CCG and merge those data sources into one

```{r import.CCG.metrics, eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
##############################################
################### POPULATION ###############
##############################################

pop_by_CCG <- fread(paste0(rawdatadir,"Other data/Mid-year population estimates/CCG/pop_by_LSOA_agg.csv"), header=TRUE, sep=",", check.names=T)

##############################################
################### DEPRIVATION ##############
##############################################

IMD_by_CCG <- fread(paste0(rawdatadir,"Other data/Lookups/CCG_to_IMD_2020_lookup.csv"), header=TRUE, sep=",", check.names=T)

##########################################################################
################### NUMBER OF SHIELDERS BY CCG: England ##################
##########################################################################

#March 2021
SPL_by_CCG_20March <- fread(paste0(rawdatadir,"SPL/England/March 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-03-20.csv"), header=TRUE, sep=",", check.names=T)
SPL_by_CCG_12March <- fread(paste0(rawdatadir,"SPL/England/March 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-03-12.csv"), header=TRUE, sep=",", check.names=T)
SPL_by_CCG_06March <- fread(paste0(rawdatadir,"SPL/England/March 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-03-06.csv"), header=TRUE, sep=",", check.names=T)

#February 2021
SPL_by_CCG_26Feb <- fread(paste0(rawdatadir,"SPL/England/February 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-02-26.csv"), header=TRUE, sep=",", check.names=T)
SPL_by_CCG_22Feb <- fread(paste0(rawdatadir,"SPL/England/February 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-02-22.csv"), header=TRUE, sep=",", check.names=T)

#15 Feb is the first dataset after QCOVID
SPL_by_CCG_15Feb <- fread(paste0(rawdatadir,"SPL/England/February 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-02-15.csv"), header=TRUE, sep=",", check.names=T)

#10 Feb is the last dataset before QCOVID  
SPL_by_CCG_10Feb <- fread(paste0(rawdatadir,"SPL/England/February 2021/CCG/Coronavirus Shielded Patient List, England - Open Data with CMO DG - CCG - 2021-02-10.csv"), header=TRUE, sep=",", check.names=T)

#Append datasets from different dates
SPL_by_CCG <- plyr::rbind.fill(SPL_by_CCG_20March,
                              SPL_by_CCG_12March,
                              SPL_by_CCG_06March,
                              SPL_by_CCG_26Feb,
                              SPL_by_CCG_22Feb,
                              SPL_by_CCG_15Feb,
                              SPL_by_CCG_10Feb)

rm(SPL_by_CCG_20March,SPL_by_CCG_12March,SPL_by_CCG_06March,SPL_by_CCG_26Feb,SPL_by_CCG_22Feb,SPL_by_CCG_15Feb,SPL_by_CCG_10Feb)

############ Merge in population numbers

length(unique(SPL_by_CCG$CCG.Code))

#Merge in population numbers

nrow(SPL_by_CCG)
SPL_by_CCG <- left_join(SPL_by_CCG,pop_by_CCG,by=c("CCG.Code"="CCG20CDH"))
nrow(SPL_by_CCG)

#Manually fix pop19 for England

SPL_by_CCG$pop19[which(SPL_by_CCG$CCG.Name=="ENGLAND")] <- sum(pop_by_CCG$pop19)

############ Save

fwrite(SPL_by_CCG, file = here::here("Clean data","SPL_by_CCG.csv"), sep = ",")
```

## Clean merged data

```{r CCG.cleaning, eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
############ Reduced dataset for all shielders

SPL_by_CCG_All <- dplyr::filter(SPL_by_CCG,Breakdown.Field=="ALL")

#Compute % shielding in each CCG

SPL_by_CCG_All <- SPL_by_CCG_All %>%
  mutate(.,Patient.Count=as.numeric(Patient.Count)) %>%
  mutate(.,Shielders_pct=Patient.Count/pop19*100) %>%
  arrange(.,Extract.Date,desc(Shielders_pct)) %>% as.data.table()

############ Find how many added by QCOVID and merge into 'ALL' dataset

QCOVID_totals <- SPL_by_CCG %>%
  filter(.,Breakdown.Field=="Disease Group",
         Breakdown.Value=="Added by COVID-19 Population Risk Assessment") %>%
  select(.,Extract.Date,CCG.Code,Patient.Count) %>%
  dplyr::rename(.,Patient.Count.QCOVID=Patient.Count)

SPL_by_CCG_All <- left_join(SPL_by_CCG_All,QCOVID_totals,by=c("Extract.Date","CCG.Code")) %>%
  mutate(.,Patient.Count.QCOVID=as.numeric(Patient.Count.QCOVID),
         Patient.Count.nonQCOVID=ifelse(is.na(Patient.Count.QCOVID),Patient.Count,Patient.Count-Patient.Count.QCOVID),
         Shielders_from_QCOVID_pct=Patient.Count.QCOVID/Patient.Count*100)

rm(QCOVID_totals)

############ Merge in deprivation

nrow(SPL_by_CCG_All)
SPL_by_CCG_All <- left_join(SPL_by_CCG_All,select(IMD_by_CCG,CCG20CDH,IMD19_decile,IMD19_quintile),by=c("CCG.Code"="CCG20CDH"))
nrow(SPL_by_CCG_All)

############ Save

fwrite(SPL_by_CCG_All, file = here::here("Clean data","SPL_by_CCG_All.csv"), sep = ",")

############ Before/After dataset - long (Feb 10 vs. March 20)

before_after_data_long <- SPL_by_CCG_All %>%
  filter(.,Extract.Date %in% c("10/02/2021","20/03/2021")) %>%
  select(.,Extract.Date,CCG.Code,CCG.Name,Patient.Count,Patient.Count.QCOVID,
         Shielders_pct,Shielders_from_QCOVID_pct,IMD19_decile,IMD19_quintile,
         pop19) %>%
  mutate(.,before_after=ifelse(Extract.Date=="20/03/2021","after","before")) %>%
  select(.,-"Extract.Date")

fwrite(before_after_data_long, file = here::here("Clean data","CCG_before_after_data_long.csv"), sep = ",")

############ Before/After dataset - wide (Feb 10 vs. March 20)

before_after_data <- SPL_by_CCG_All %>%
  filter(.,Extract.Date %in% c("10/02/2021","20/03/2021")) %>%
  select(.,Extract.Date,CCG.Code,CCG.Name,Patient.Count,Patient.Count.QCOVID,
         Shielders_pct,Shielders_from_QCOVID_pct,IMD19_decile,IMD19_quintile,
         pop19) %>%
  mutate(.,before_after=ifelse(Extract.Date=="20/03/2021","after","before")) %>%
  select(.,-"Extract.Date") %>%
  pivot_wider(.,
              names_from = before_after,
              names_sep = "_",
              values_from = c(Patient.Count,Patient.Count.QCOVID,
                              Shielders_pct,Shielders_from_QCOVID_pct)) %>%
  mutate(.,Patient.Count_diff=Patient.Count_after-Patient.Count_before,
         Patient.Count_diff=Patient.Count_after-Patient.Count_before,
         Shielders_pct_diff=Shielders_pct_after-Shielders_pct_before)

fwrite(before_after_data, file = here::here("Clean data","CCG_before_after_data.csv"), sep = ",")
```

## Descriptive statistics used in final document

## People in England asked to shield before/after QCOVID

```{r CCG.desc.stats.0, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
fread_folder(paste0(gitdir,"/Clean data/"), extension = "CSV")

CCG_before_after_data %>%
  filter(.,CCG.Name=="ENGLAND") %>%
  select(.,CCG.Name,Shielders_pct_before,Shielders_pct_after) %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling()

data.QCOVID.chart <- CCG_before_after_data_long %>%
  select(.,CCG.Name,before_after,Shielders_pct,IMD19_quintile) %>%
  filter(.,(!is.na(Shielders_pct))&(!is.na(IMD19_quintile))) %>%
  mutate(., IMD19_quintile=ifelse(IMD19_quintile=="1","1 (most deprived)",IMD19_quintile)) %>%
  mutate(., IMD19_quintile=ifelse(IMD19_quintile=="5","5 (least deprived)",IMD19_quintile)) %>%
  arrange(.,CCG.Name,before_after)
fwrite(data.QCOVID.chart, file = paste0(briefonedrivedir,"data.QCOVID.chart.csv"), sep = ",")
```

## CCGs with most/least people shielding and those with highest change attributed to QCOVID

```{r CCG.desc.stats.1, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
#CCGS with highest % shielding after QCOVID
CCG_before_after_data %>%
  slice_max(.,Shielders_from_QCOVID_pct_after,n=5) %>%
  select(.,CCG.Name,Shielders_pct_before,Shielders_pct_after,Shielders_from_QCOVID_pct_after) %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling()

#CCGS with lowest % shielding after QCOVID
CCG_before_after_data %>%
  slice_min(.,Shielders_from_QCOVID_pct_after,n=5) %>%
  select(.,CCG.Name,Shielders_pct_before,Shielders_pct_after,Shielders_from_QCOVID_pct_after) %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling()

#CCGS with highest change in % shielding after QCOVID
CCG_before_after_data %>%
  slice_max(.,Shielders_pct_diff,n=5)  %>%
  select(.,CCG.Name,Shielders_pct_diff) %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling()
```

## Aggregated % shielding and std. dev. of % shielding by IMD quintile

```{r CCG.desc.stats.2, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
SPL_by_CCG_All %>%
  filter(.,(Extract.Date %in% c("10/02/2021","20/03/2021")),CCG.Name!="ENGLAND") %>%
  select(.,Extract.Date,CCG.Name,Patient.Count,Shielders_pct,pop19,IMD19_quintile) %>%
  group_by(Extract.Date,IMD19_quintile) %>%
  summarise(.,pop19=sum(pop19,na.rm=TRUE),
            Patient.Count=sum(Patient.Count,na.rm=TRUE),
            std.dev.pct.shielding=sd(Shielders_pct)) %>%
  mutate(.,Shielders_pct_agg=Patient.Count/pop19*100) %>%
  ungroup() %>%
  arrange(.,IMD19_quintile,Extract.Date) %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling()
```


